Telemetry Module for Solar Vehicle

### 2021-06-17 documentation/FinalPaper/paper.tex
Updates from Overleaf

\usepackage[pdftex]{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{float}
\usepackage{comment}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[style=ieee]{biblatex}
\addbibresource{paper.bib}
\usepackage{url}
\usepackage[linkcolor = blue]{hyperref}
\newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Abstract
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
ROAST, telemetry, CAN
\end{IEEEkeywords}

@@

@@

This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project. Thereafter, the design... \todo{Skal laves når strukturen er slået fast}

@@

\section{Design}
The overall system consists of three separate modules; the Solar Car Unit (SCU) which is connected to the CAN bus and communicates with the support vehicle unit via a wireless RF connection. The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial UART connection. The GUI program running on the laptop broadcasts the CAN bus message stream on a UDP port on the local network. A block diagram for the system is outlined in Fig. \ref{fig:schematic}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/SystemSchematic.pdf}
    \caption{Overview of the system's components in block diagram form.}
    \label{fig:schematic}
\end{figure}
%

@@

\subsection{Microcontrollers} % Steffan
We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU respectively. The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data. Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.

\subsection{CAN bus}
The CAN network in the SCU enables the all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
    \caption{Structure of a CAN frame}
    \label{fig:CANframe}
\end{figure}

@@

The Teensy 3.6 used in the solar car features a CAN controller handling the communication with the CAN bus. In order to interface with the physical CAN bus though, a transceiver is needed. Here the MCP2551 was chosen, which maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
% MCP
\subsection{RF network} % Steffan
The system uses two nRF24L01+ transceivers to provide a bidirectional connection between the Solar Car and Support Vehicle over the 2.4 GHz ISM band. The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data transfer rates of up to 2 Mbit/s.

An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}. Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Fig. \ref{fig:shockburst}. Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically and therefore fall beyond the scope of this project.

@@

We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot. Furthermore, this solution scales very easily for larger storage capacity, with new SD card sizes exceeding 1 TiB.

@@

As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.

\subsection{Other Libraries}

In order to interface with the components presented in the previous section, a set of libraries were used.

The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards. The library has been important for the connection between the CAN network and the ports on the Teensy board. We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD. The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.

% ACAN.h
% RF24.h
%

@@

\todo{make clearer distinction between message types and commands. Several commands can share the same message type}

The first byte, the command, always identifies the message type and indicates how the remaining 15 bytes should be interpreted. This eases the decoding of messages on the receiving end and makes for a expandable framework which has enough room for future expansions.

As an example of a specific message type, figure \ref{fig:messageTypes} shows a message at the bottom with the command \texttt{BROADCAST\_CAN} which is encoded by \texttt{0x01}. This message carries a CAN message as a payload and commands the solar car subsystem to inject the message into the solar car CAN bus. Even though some message types are only sent one direction between the telemetry modules, a unified design was chosen due to the inherent simplicity.

As an extra feature not included in our original specification, we have made an encryption protocol for our telemetry module. The encryption is based on the RSA algorithm, which essentially generates an encryption key based on the modulus and totient function of two prime numbers. The pairings of prime numbers suited for the encryption protocol can be seen below in Table \ref{fig:prime_tabel}. When a message is encrypted, each of the 16 bytes is extended to a 16-bit unsigned integer to avoid overflow problems. Once transmitted to the receiving node, the message is then decrypted and converted back to unsigned 8-bit type. This essentially doubles the size of the data payloads being sent in our RF transmission, which is the downside of our encryption solution.

@@

% message protcol
% command system
% security -> encryption

\section{\todo{Final product}}
\subsection{Solar Car}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{documentation/images/SCU_CircuitDiagram.pdf}
    \caption{The circuit diagram for the SCU.}
    \label{fig:SCU_circuit}
\end{figure}
% threads 
\subsection{Support Vehicle}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{documentation/images/SVU_CircuitDiagram.pdf}
    \caption{The circuit diagram for the SVU.}
    \label{fig:SCU_circuit}
\end{figure}
% threads
\subsection{\todo{Graphical User Interface}}
% GUI

\section{Testing and Validation} % Victor
As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design. We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.

We performed software validation as we went along with the software coding, to make sure that the actual behaviour of the code was as the expected behaviour. This was done through the testing environment in PlatformIO. We used this to validate our code design with unit tests, testing both edge cases and random cases. We chose only to validate the code and functions we made ourselves, and not those that was imported from external libraries.

@@

%Finally, we were able to test all three modules

@@

% throughput limitations
% variable length messages would reduce overhead -> but most messages contain CAN and are optimized

>>>>

\usepackage[pdftex]{graphicx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{float}
\usepackage{comment}
\usepackage{movie15}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[style=ieee]{biblatex}
\addbibresource{paper.bib}
\usepackage{url}
\usepackage[linkcolor = blue]{hyperref}
\newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
This paper presents the telemetry project for the DTU ROAST Solar car. The development of the module is done through design sprints, where hardware and software is gradually developed along side each other. First a Hardware prototype on breadboards, then the software to run the modules satisfactorily was developed and finally the final product was made when the hardware was soldered to perf boards.\\
The software developed for the module provided the module to have two-way communication with CAN message structure, encryption of the data during transmission between the module in the solar car and a module in a support car, the ability to give the solar car module commands from the support car and streaming received messages in the support car to Matlab.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
ROAST, telemetry, CAN, RSA encryption, Matlab
\end{IEEEkeywords}

@@

\begin{figure}[b]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/SolarChallengeLogo.pdf}
    \caption{Bridgestone World Solar Challenge logo \cite{wsc}.}
    \label{fig:solar}
\end{figure}

@@

This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project. Thereafter, the overall system design is presented and the following paragraphs discuss the hardware components and software used for these. After this comes a paragraph discussing the implementation of the communication between the two modules, both internal and during transmission. This concludes with a section of the final product, which is followed up by our method method for testing and validating our solution, as well as discussion and conclusion of the solution.

@@

\section{Design}
The overall system consists of three separate modules; the Solar Car Unit (SCU) which is connected to the CAN bus and communicates with the support vehicle unit via a wireless RF connection. The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial connection. The GUI program running on the laptop broadcasts the CAN bus message stream on the local network. A block diagram for the system is outlined in Fig. \ref{fig:schematic}. 
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{documentation/images/SystemSchematic.pdf}
    \caption{Overview of the system's components in block diagram form.}
    \label{fig:schematic}
\end{figure}

@@

\subsection{Microcontrollers} % Steffan
We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU, respectively. The Teensy line-up has been used as the main microcontroller platform in most ecocar projects and we decided to use them as well after considering the performance of newer Teensy boards. The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data. Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.

\subsection{CAN bus}
The CAN network in the SCU enables all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
    \caption{Structure of a CAN frame.}
    \label{fig:CANframe}
\end{figure}

@@

The Teensy 3.6 used in the solar car features a CAN controller handling message assembly and receiving from the CAN bus. In order to interface with the physical CAN bus though, a transceiver is needed. Here the MCP2551 was chosen since it has been used extensively in other ecocar projects. The component maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.

\subsection{RF network} % Steffan
Two radio-frequency (RF) transceivers are required to provide a stable connection between the SCU and SVU at longer distances. There are many options for radio modules supporting long range communication at the expense of achievable data rate. We chose a trade-off between both by selecting the nRF24L01+ operating on the 2.4 GHz ISM band. The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m and data transfer rates of up to 2 Mbit/s.

An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}. Messages are transmitted in packages, structured in flexible byte segments, as shown in Fig. \ref{fig:shockburst}. Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically and therefore fall beyond the scope of this project.

@@

We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot. Furthermore, this solution scales very well and easily in terms storage capacity, with new SD card sizes exceeding 1 TiB.

@@

As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported on both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.

\subsection{Graphical User Interface} % Tjark

The GUI included in the telemetry system serves a proof-of-concept purpose. Therefore, scala-swing \cite{scala-swing} was chosen as a framework, due to the effectiveness of the scala \cite{scala} language and the high level of abstraction used when describing a user interface in scala-swing. Furthermore, a GUI implementation in scala can run on any system with a java virtual machine and is therefore cross-platform compatible.

In order to send and receive data from the SVU via serial connection, the jSerialComm \cite{jSerialComm} library was used.

\subsection{Data Processing} % Steffan
Due to the wishes of DTU ROAST to use the telemetry data in applications of predictive modelling and performance optimization we decided to implement a simple UDP network stream hosted on the Support Vehicle laptop. This would allow all devices on the local network to monitor the data being sent over the serial connection for use in further data processing. For a proof-of-concept case we created a simple Matlab script to sample CAN messages from the UDP port and print them formatted to the terminal.

\subsection{Other Libraries}
In order to interface with the components presented in the previous section, a set of libraries were included in the project. These include a CAN network driver to support use of the Teensy 3.6's integrated CAN controller \cite{ACAN}, an RF OSI layer providing seamless integration with the nRF24L01+ modules, and an SD card driver to control data logging in the black box memory.

%In order to interface with the components presented in the previous section, a set of libraries were used.

%The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards. The library has been important for the connection between the CAN network and the ports on the Teensy board. We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD. The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s.

@@

The system is built around \textit{commands} and \textit{message types}. All message types begin with a command byte. The remaining 15 bytes are interpreted based on the message type. As an example, a \texttt{CanTelemetryMsg} message type is shown in figure \ref{fig:messageTypes}.

One message type can be connected to different commands. For instance, the \texttt{CanTelemetryMsg} message type can be used to send data from the SCU to the SVU as part of the data stream setting the command to \texttt{RECEIVED\_CAN} = \texttt{0x00} or it can be used to inject a CAN message from the SVU into the solar car's CAN bus setting the command to \texttt{BROADCAST\_CAN} = \texttt{0x01}.

When decoding messages, only the first byte needs to be considered and it will unambiguously decide how the rest of the message should be interpreted. This eases the decoding of messages on the receiving end and makes for a expandable framework which has enough room for future expansions.

Even though some message types are only sent one direction between the telemetry modules, a unified design was chosen due to the inherent simplicity.

As an extra feature not included in our original specification, we have made an encryption protocol for our telemetry module. The encryption is based on the RSA algorithm, which essentially generates an encryption key based on the modulus and totient function of two prime numbers. The pairings of prime numbers suited for the encryption protocol can be seen in Fig. \ref{fig:prime_tabel}.

When a message is encrypted, each of the 16 bytes is extended to a 16-bit unsigned integer to avoid overflow problems increasing the total message size to 32 bytes. Once transmitted to the receiving node, the message is then decrypted and the overflow protection bytes can be removed leaving us with 16 bytes again. This essentially doubles the size of the data payloads being sent in our RF transmission, which is the downside of our encryption solution.

@@

\section{\todo{Final product}}
\subsection{Solar Car Unit}

% actual circuit
% - powered through molex connector 12V
% - converted to 5V and teensy internal voltage regulator to 3.3
% - Teensy has RTC which can keep time when powered by an external 2.9V battery

% on the software side
% - threads:
%  -

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{documentation/images/SCU_CircuitDiagram.pdf}
    \caption{The circuit diagram for the SCU.}
    \label{fig:SCU_circuit}
\end{figure}
% threads 
\subsection{Support Vehicle Unit}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{documentation/images/SVU_CircuitDiagram.pdf}
    \caption{The circuit diagram for the SVU.}
    \label{fig:SCU_circuit}
\end{figure}
% threads
\subsection{\todo{Graphical User Interface}}
% GUI

\section{Verification and Validation} % Victor
As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design. We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.

We performed software validation concurrently with writing our code to make sure that the actual behaviour of the program matched the expected behaviour. This was done through the testing environment in PlatformIO. We used this to validate our code design with unit tests, testing both edge cases and random cases. We chose only to validate the code and functions we made ourselves, and not those that was imported from external libraries.

@@

@@

% throughput limitations
% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
% UDP works only one way for now. If automated control from matlab should be possible, udp listeners should be able to send as well
